### 2015-04-19 src/5_3_Proposal.tex
ultimos cambios

In order to formulate this algorithm we need to define first the
following matrices which considers only the last $L$ instances:

@@

The RR solution $\mathbf{X}(t)$ using the sliding window matrices defined above
is:

\begin{equation}
\label{eq:optsolSLAAR}
%\mathbf{X}(t)&=&()^{-1}{\bf A}(t)^\top{\bf B}(t) \\
\mathbf{X}(t)=\mathbf{S}(t)^{-1} \mathbf{W}(t) \, ,
\end{equation}

@@

\begin{eqnarray*}
\mathbf{S}(t) =&{\bf A}(t)^\top{\bf A}(t)+ \gamma \mathbb{I} &=
\sum_{i=0}^L \mathbf{a}_{t-i}\mathbf{a}_{t-i}^\top + \gamma \mathbb{I}
\label{eq:S} \\
\mathbf{W}(t) =& {\bf A}(t)^\top{\bf B}(t) &= 
\sum_{i=0}^L \mathbf{a}_{t-i}\mathbf{b}_{t-i} 
\label{eq:W}
\end{eqnarray*}

@@

Therefore Sherman-Morrison-Woodbury formula can be used to reduce inverse matrix
calculations.

The algorithm~\ref{alg:proposal} shows our online VECM using three different
methods for geting model parameters: OLS, RR, AAR.

@@

\begin{itemize}
\item The function \texttt{getJohansen} returns cointegration vectors given by the
Johansen method considering the trace statistic test at 95\% level of
significance. This procedure is called at the first step of the algorithm and
every time in-sample MAPE ($\mathbf{e}$) exceeds threshold error defined (mean\_error).
\item The function \texttt{vecMatrix} returns VECM
matrices $\mathbf{A},\mathbf{B}$ shown in equation~(\ref{eq:vecmatrix}). This
method is only required at the first step.
\item The function \texttt{vecMatrixOnline} returns new rows of $\mathbf{A}$ and
$\mathbf{B}$ give new input data $\mathbf{y}_t$.
\item solver (see algorithm \ref{alg:solver}) is an object which allows online
calculation saving matrices $\mathbf{S}$ and $\mathbf{W}$ (see
equations~\ref{eq:S} and \ref{eq:W}) for methods RR and AAR. This method is
initialized in the first step and only if new cointegration are required. This
method also returns in-sample MAPE saved in $\mathbf{e}$. 
\item The function \texttt{vecMatrixUpdate} updates matrix $\mathbf{A}$ when new
cointegration vectors are required (matrix $\mathbf{B}$ is not affected). Model
parameters $\mathbf{X}$ is also updated later.
\item Out-of-sample outputs are saved in the variables $\Delta
\mathbf{Y}_{\text{true}}$ and $\Delta \mathbf{Y}_{\text{pred}}$ which are used
to obtain Out-of-sample MAPE at the end.
%\item The number of cointegration vectors is set as a parameter (aun no lo
%coloco en el algoritmo).
%\item In order to set VECM parameter: $L$ and $p$, we use the Akaike Information
%Criterion (AIC). RR parameter $\lambda$ was done by cross-validation.
%\item Obtain VECM parameter $\mathbf{X}(t)$ in equation \ref{eq:optsolSLAAR} requires
%to calculate the inverse of matrix $\mathbf{S}(t+1)$ which is an expensive
%routine.However, since we already know $\mathbf{S}(t)^{-1}$ we can use
%Sherman-Morrison-Woodbury twice~\ref{eq:SMW}.
\end{itemize}

>>>>

VECM($p$) for $l$ cointegrated variables has the following form:

\begin{equation}\label{eq:vecfull}
\Delta\mathbf{y}_t 
= \boldsymbol{\alpha\beta}^\top\mathbf{y}_{t-1} 
  + \sum_{i=1}^{p-1}\boldsymbol{\Phi}_i^*\,\Delta\mathbf{y}_{t-i}
  + \mathbf{c} + \boldsymbol{\epsilon}_t \qquad \forall t = p+1,\dots,N.
\end{equation}

Transposing each equation of the system (\ref{eq:vecfull}) we can write
the VECM($p$) model in block-matrix form as:
\begin{equation}\label{eq:vareq}
\mathbf{B} = 
\mathbf{A} \mathbf{X} + 
\mathbf{E} \, , 
\end{equation}
%
\noindent where $\mathbf{B}$ dimension is $((N-p)\times l)$, $\mathbf{A}$
dimension is $((N-p)\times(r+(p-1)l +1))$, $\mathbf{X}$ dimension is $((r+(p-1)l
+1)\times l)$, $\mathbf{E}$ dimension is $((N-p)\times l)$ and $r$ is the number
of cointegration vectors:
%
\begin{alignat}{3}
\mathbf{B}
&= \begin{bmatrix}
   \Delta\mathbf{y}_{p+1}^\top \\
   \Delta\mathbf{y}_{p+2}^\top \\
   \vdots \\
   \Delta\mathbf{y}_N^\top
   \end{bmatrix}
&\quad
\mathbf{X}
&= \begin{bmatrix}
   \boldsymbol{\alpha}^\top \\
   \boldsymbol{\Phi}_1^{*\top} \\
   \boldsymbol{\Phi}_2^{*\top} \\
   \vdots \\
   \boldsymbol{\Phi}_{p-1}^{*\top} \\
   \mathbf{c}^\top
   \end{bmatrix}
&\quad
\mathbf{E}
&= \begin{bmatrix}
   \boldsymbol{\epsilon}_{p+1}^\top \\
   \boldsymbol{\epsilon}_{p+2}^\top \\
   \vdots \\
   \boldsymbol{\epsilon}_N^\top \\
   \end{bmatrix}
\end{alignat}
\noindent and 
\begin{align}
\mathbf{A} 
&= \begin{pmat}[{....|}]
   \mathbf{y}_p^\top \boldsymbol{\beta} & \Delta \mathbf{y}_p^\top & \Delta\mathbf{y}_{p-1}^\top & \dots 
                    & \Delta\mathbf{y}_2^\top & 1 \cr
   \mathbf{y}_{p+1}^\top  \boldsymbol{\beta} &\Delta\mathbf{y}_{p+1}^\top & \Delta\mathbf{y}_p^\top & \dots
                       & \Delta\mathbf{y}_3^\top & 1 \cr
   \vdots & \vdots & \vdots & \ddots & \vdots & \vdots \cr
   \mathbf{y}_{N-1}^\top  \boldsymbol{\beta} &\Delta\mathbf{y}_{N-1}^\top & \Delta\mathbf{y}_{N-2}^\top & \dots 
                       & \Delta\mathbf{y}_{N-p-1}^\top & 1 \cr
   \end{pmat}\, .
\end{align}
%Taking into account the error term $\mathbf{E}$, equation~(\ref{eq:vareq}) 
%can be solved with respect to $\mathbf{X}$ using the ordinary least
%squares estimation.
However, the number of rows of matrix $\mathbf{A}$ increases with the number of
observations $\mathbf{y}_t$. Our proposal considers only the most recent $L$
rows of matrices $\mathbf{A}$ and $\mathbf{B}$ defined as  $\mathbf{A}(t)$ and
$\mathbf{B}(t)$:

@@

\noindent so that VECM parameters $\mathbf{X}(t)$ are found using the following
equation:

\begin{equation}
\mathbf{B}(t) = \mathbf{A}(t) \mathbf{X}(t) + \mathbf{E}(t)\\
\end{equation}

The RR solution $\mathbf{\hat{X}}(t)$ using the sliding window matrices defined above
is:

\begin{equation}
\label{eq:oproblem}
\mathbf{\hat{X}}(t)=\mathbf{S}(t)^{-1} \mathbf{W}(t) \, ,
\end{equation}

@@

\begin{eqnarray*}
\mathbf{S}(t) =&{\bf A}(t)^\top{\bf A}(t)+ \lambda \mathbb{I} &=
\sum_{i=0}^L \mathbf{a}_{t-i}\mathbf{a}_{t-i}^\top + \lambda \mathbb{I}
\label{eq:S} \\
\mathbf{W}(t) =& {\bf A}(t)^\top{\bf B}(t) &= 
\sum_{i=0}^L \mathbf{a}_{t-i}\mathbf{b}_{t-i} 
\label{eq:W}
\end{eqnarray*}

@@

%Agregar solo si se decide agregar Sherman-Morrison
%Therefore the Sherman-Morrison-Woodbury formula can be used to reduce inverse matrix
%calculations.

The algorithm~\ref{alg:proposal} shows OVECM using three different
methods for geting model parameters: OLS, RR, AAR which leads to three different
algorithms OVECM-OLS, OVECM-RR, OVECM-AAR:

@@

\begin{itemize}
\item The function \texttt{getJohansen} returns cointegration vectors given by the
Johansen method considering the trace statistic test at 95\% level of
significance. This procedure is called at the first step of the algorithm and
every time in-sample MAPE ($\mathbf{e}$) exceeds threshold error defined (mean\_error).
\item The function \texttt{vecMatrix} returns VECM
matrices $\mathbf{A}(t),\mathbf{B}(t)$ shown in equation~(\ref{eq:notation}). This
method is only required at the first step.
\item The function \texttt{vecMatrixOnline} returns new rows $\mathbf{a}_t^\top$ and
$\mathbf{b}_t^\top$ given new input data $\mathbf{y}_t$.
\item The Solver class (see algorithm \ref{alg:solver}) obtain VECM parameters  
using RR, AAR or OLS methods.
\item The function \texttt{vecMatrixUpdate} updates matrix $\mathbf{A}$ when new
cointegration vectors are required (matrix $\mathbf{B}$ is not affected). Model
parameters $\mathbf{X}$ is also updated later.
%\item The number of cointegration vectors is set as a parameter (aun no lo
%coloco en el algoritmo).
%\item In order to set VECM parameter: $L$ and $p$, we use the Akaike Information
%Criterion (AIC). RR parameter $\lambda$ was done by cross-validation.
%\item Obtain VECM parameter $\mathbf{X}(t)$ in equation \ref{eq:optsolSLAAR} requires
%to calculate the inverse of matrix $\mathbf{S}(t+1)$ which is an expensive
%routine.However, since we already know $\mathbf{S}(t)^{-1}$ we can use
%Sherman-Morrison-Woodbury twice~\ref{eq:SMW}.
\end{itemize}
